Analysis of NLA17 Dissolved Gas Indicator

### 2019-12-12 scripts/dgIndicatorAnalysis.Rmd
separating data prep from analysis

## Getting data and filtering
The dissolved gas data were currated the 'NLA' dissolved gas project in RStudio.  The code can be found at a private github repository (https://github.com/USEPA/NLA).  After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2019-10-25.txt.  A copy of this data file is stored in documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).  When synced to a local computer, the documents library can be read directly with R, after identifying the local directory path.

```{r}
# Read data file
dg <- read.table(file = paste0(localPath, 
                               "/Environmental Protection Agency (EPA)/",
                               "ORD NLA17 Dissolved Gas - Documents/",
                               "inputData/nla17gasDataAggregated_2019-11-12.txt"),
                 header = TRUE, sep = "\t", as.is = TRUE) %>%
  # unit conversion
  mutate(dissolved.ch4 = dissolved.ch4 * 1000000, # mol/L -> umol/L
         dissolved.co2 = dissolved.co2 * 1000000, # mol/L -> umol/L,
         dissolved.n2o = dissolved.n2o * 1000000000, # mol/L -> nmol/L
         # add source/sink column
         co2.src.snk = ifelse(co2.sat.ratio > 1, "source", "sink"),
         ch4.src.snk = ifelse(ch4.sat.ratio > 1, "source", "sink"),
         n2o.src.snk = ifelse(n2o.sat.ratio > 1, "source", "sink"),
         # add state
         state = substr(site.id, 7, 8))

```
Concentrations of CO~2~, CH~4~, and N~2~O in ambient air over the waterbodies were measured and used in the dissolved gas concentrations.  These data are not needed in subsequent analysis and will be removed from the dataframe.
```{r}
dg <- dg %>% filter(sample.source == "DG") # select dissolved gas 'DG' samples
```
The dataframe also contains observations from site in Alaska.  These sites were not an official part of the NLA and will not be included in this analysis.
```{r}
dg <- dg %>% filter(!(grepl("AK", site.id))) # exlclude alaska
```

>>>>

## Getting data and filtering
The dissolved gas data were currated the 'NLA' dissolved gas project in RStudio.  The code can be found at a private github repository (https://github.com/USEPA/NLA).  After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2019-10-25.txt.  A copy of this data file is stored in documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).  This file was futher manipulated in "C:\Users\JBEAULIE\OneDrive - Environmental Protection Agency (EPA)\gitRepository\DissolvedGasNla\scripts\dataMunge.Rmd" and written out as an .RData object which is imported below.

```{r}
# Read data file
load(paste0(localPath, 
            "/Environmental Protection Agency (EPA)/",
            "ORD NLA17 Dissolved Gas - Documents/",
            "inputData/dg.RData"))
